\documentclass{article}
\pagenumbering{gobble}

\begin{document}
	
	\vspace*{-1in}
	
	\begin{center}
	\textsc{
		Travis Scholl $\vert$ {\tt tscholl2@uw.edu} \\
		Department of Mathematics \\
		University of Washington \\
		Seattle, WA 98195
	}
	\end{center}
	
	\section{Profile}
	
	Ph.D. student in mathematical cryptography with 8+ years of programming experience in various languages including Python, Javascript, and Go. Spent 2+ years working as a full-stack developer designing and building an online math homework system for the University of Washington. Has the initiative to work well independently, and the communication skill to succeed in collaborative environments.
	
	\section{Skills}
	
	\begin{itemize} 
		\item 5+ years of Python, Javascript, 4+ years of Go, 2+ years of Typescript, Coffeescript, SQL.
		\item 5+ years of graduate research in mathematical cryptography.
		\item 3+ years of contributing to open source projects.
	\end{itemize}

	\section{Professional Experience}
	
	\begin{itemize}  
		\item 2013 - 2018: Graduate Teaching Assistant/Instructor, University of Washington
			\subitem Teach calculus courses, grade exams, lead discussion sections
		\item 2015 - 2017: Full Stack Developer, University of Washington
			\subitem Design, build, and maintain an online math homework system.
		\item 2015 - 2016: Contract Web Designer, Virginia Tech
			\subitem Review, update, and manage code for several dynamic webpages.
	\end{itemize}
	
	\section{Education}
	
	\begin{itemize}  
		\item 2018 (Expected): Ph.D. in Mathematics, University of Washington.
		\item 2013: B.S. in Mathematics, University of Oregon.
	\end{itemize}
	
	\section{Personal Skills}
	
	\begin{description}
		\item[Initiative:] Can independently learn new material as needed for projects.
		\item[Critical thinking:] Ability to break down large problems into manageable parts.
		\item[Communication:] Clearly convey technical solutions to both specialized and general audiences.
	\end{description}

	
\end{document}